Undirected Graph Exploration with $\Theta(\log\log n)$ Pebbles

### 2016-05-09 presentation/main.tex
finished simulation

\usepackage[utf8]{inputenc}
\usepackage{epigraph}
\usepackage{tikz}
\usepackage{csquotes}
\usepackage{xcolor}

@@

\section{Graph traversal}
\begin{frame}
  \frametitle{Exploring pebble machine}
  \begin{theorem}[Reingold]
    There is an $\mathcal{O}(n)$-space algorithm producing a universal
    exploration sequence for any regular graph on $n$ vertices.
  \end{theorem}
  \begin{center}
    \begin{tikzpicture}
      \node (d1) {};
      \node[below=2 of d1] (d2) {};
      \draw[decorate,decoration={snake,amplitude=0.3cm, segment length=1cm},->]
        (d1) to (d2);
    \end{tikzpicture}
  \end{center}
  \begin{theorem}
    There exists a $(\mathcal{O}(1), 0, \mathcal{O}(\log n))$-pebble machine
    that moves along a closed walk and either explores the graph or visits at
    least $n$ distinct vertices, for any graph with bounded degree.
  \end{theorem}
\end{frame}
\begin{frame}
  \frametitle{Closed walk}
  \begin{lemma}
    An agent following an exploration sequence of the form
    $(e_{1},\dots,e_{k})^{\ast}$ moves along a closed walk.
  \end{lemma}
  \begin{itemize}
    \item Examine triples $(v,\ell,i)$:
      \begin{itemize}
        \item $v$: current vertex
        \item $\ell$: port agent entered $v$ through
        \item $i$: next step of the exploration sequence
      \end{itemize}
    \item Deterministically $(v,\ell,i) \leadsto (v',\ell',i')$
    \item If $t_{1} = (v',\ell',i') \leadsto (v,\ell,i)$ and
      $t_{2} = (v'',\ell'',i'') \leadsto (v,\ell,i)$: $t_{1} = t_{2}$
    \item Only finite triples: projection to vertices yields closed walk
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Universal exploration of $3$-regular graphs}
  \begin{lemma}
    There exists an $\mathcal{O}(\log n)$-space algorithm producing a universal
    exploration sequence $(e_{1},\dots,e_{k})^{\ast}$ for any $3$-regular graph
    with at most $n$ vertices.
  \end{lemma}
  \begin{itemize}
    \item Let $\mathcal{M}$ be the Turing machine from Reingold's Theorem
    \item The exploration sequence computed by $\mathcal{M}$ is universial for
      3-regular graphs with at most $n$ vertices
      \alt<1-1>{
        \begin{itemize}
          \item Assume $3$-regular $G_{0}$ with $n_{0} < n$ vertices is not
            explored
          \item Then there is a vertex that is not traversed
          \item Repeatedly apply \parbox{3\baselineskip}{\resizebox{!}
            {3\baselineskip}{\input{tikz/blowup}}}
          \item Until there are $n$ vertices which contradicts Reingold's
            Theorem
        \end{itemize}
      }{
      \item Produce a periodic universal exploration sequence
        $(e_{1},\dots,e_{k})^{\ast}$
        \begin{itemize}
          \item Construct $\mathcal{M}'$ which simulates $\mathcal{M}$ for
            $3n\cdot c$ steps where $c$ is the number of configurations of $M$
          \item Output $e = (e_{1},\dots,e_{k})^{\ast}$
          \item Assume $e$ is not universal; then it loops in $V'$ with
            $|V'|<n$
          \item $\mathcal{M}$ has to enter a vertex twice with the same
            configuration
          \item Contradicts universality of the exploration sequence of
            $\mathcal{M}$
        \end{itemize}
      }
  \end{itemize}
  \pause
\end{frame}

\begin{frame}
  \frametitle{Exploring pebble machine}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item<2-> Establish 3-regularity
        \begin{itemize}
          \item Transform $v$ to $3\cdot d_{v}$ distinct vertices
        \end{itemize}
      \item<4-> Use Reingold's algorithm and take first $k = 3n\cdot c$ ($c$ number of
        configurations $M_{A}$)
    \end{itemize}
    \column{0.5\textwidth}
    \uncover<3->{\resizebox{\textwidth}{!}{\input{tikz/regsim}}}
  \end{columns}
\end{frame}

\section{Simulation}
\begin{frame}
  \frametitle{Tape simulation via pebbles}
  \begin{itemize}
    \item Simulate tape of size $m$ by pebbles:
      \begin{itemize}
        \item<3-> Separate into memory blocks of size $m_{1}$
        \item<4-> Represent every block by one of $\left\{p_{0},\dots,p_{\frac{m}{m_{1}}-1}\right\}$ pebbles
      \end{itemize}
  \end{itemize}
  \uncover<2->{\input{tikz/tapesimint}}
\end{frame}

@@

\begin{frame}
  \frametitle{Simulation}
  \begin{theorem}
    For every graph $G$ a $(s,p,2m)$ pebble machine $M$ can be simulated by a
    $(cs,p+c,m)$ pebble machine that simulates the walk of $M$ on $G$ or
    explores $G$.
  \end{theorem}
  \begin{itemize}
    \item There is a $(c,0,c\cdot\log n)$ pebble machine which explores
      a graph $G$ with $n$ vertices
    \item Applying above simulation $\log \log n$ times yields:
      \begin{theorem}
        For every Graph $G$ with $n$ vertices there is a
        $(c'^{\log\log n}\cdot c, \log\log n\cdot c', c)$ pebble machine
        which explores $G$.
      \end{theorem}
  \end{itemize}
\end{frame}

>>>>

\usepackage[utf8]{inputenc}
\usepackage{epigraph}
\usepackage{tikz}
\usepackage{csquotes}
\usepackage{xcolor}
\usepackage{amssymb}

@@

\section{Graph traversal}
\begin{frame}
  \frametitle{Exploring pebble machine}
  \begin{theorem}[Reingold]
    There is an $\mathcal{O}(n)$-space algorithm producing a universal
    exploration sequence for any regular graph on $n$ vertices.
  \end{theorem}
  \begin{center}
    \begin{tikzpicture}
      \node (d1) {};
      \node[below=2 of d1] (d2) {};
      \draw[decorate,decoration={snake,amplitude=0.3cm, segment length=1cm},->]
        (d1) to (d2);
    \end{tikzpicture}
  \end{center}
  \begin{theorem}
    There exists a $(\mathcal{O}(1), 0, \mathcal{O}(\log n))$-pebble machine
    that moves along a closed walk and either explores the graph or visits at
    least $n$ distinct vertices, for any graph with bounded degree.
  \end{theorem}
\end{frame}
\begin{frame}
  \frametitle{Closed walk}
  \begin{lemma}[Closed walk]
    An agent following an exploration sequence of the form
    $(e_{1},\dots,e_{k})^{\ast}$ moves along a closed walk.
  \end{lemma}
  \begin{itemize}
    \item Examine triples $(v,\ell,i)$:
      \begin{itemize}
        \item $v$: current vertex
        \item $\ell$: port agent entered $v$ through
        \item $i$: next step of the exploration sequence
      \end{itemize}
    \item Deterministically $(v,\ell,i) \leadsto (v',\ell',i')$
    \item If $t_{1} = (v',\ell',i') \leadsto (v,\ell,i)$ and
      $t_{2} = (v'',\ell'',i'') \leadsto (v,\ell,i)$: $t_{1} = t_{2}$
    \item Only finite triples: projection to vertices yields closed walk
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Universal exploration of $3$-regular graphs}
  \begin{lemma}[Periodic exploration sequence (PES-Lemma)]
    There exists an $\mathcal{O}(\log n)$-space algorithm producing a universal
    exploration sequence $(e_{1},\dots,e_{k})^{\ast}$ for any $3$-regular graph
    with at most $n$ vertices.
  \end{lemma}
  \begin{itemize}
    \item Let $\mathcal{M}$ be the Turing machine from Reingold's Theorem
    \item The exploration sequence computed by $\mathcal{M}$ is universial for
      3-regular graphs with at most $n$ vertices
      \alt<1-1>{
        \begin{itemize}
          \item Assume $3$-regular $G_{0}$ with $n_{0} < n$ vertices is not
            explored
          \item Then there is a vertex that is not traversed
          \item Repeatedly apply \parbox{3\baselineskip}{\resizebox{!}
            {3\baselineskip}{\input{tikz/blowup}}}
          \item Until there are $n$ vertices which contradicts Reingold's
            Theorem
        \end{itemize}
      }{
      \item Produce a periodic universal exploration sequence
        $(e_{1},\dots,e_{k})^{\ast}$
        \uncover<2-2>{
          \begin{itemize}
            \item Construct $\mathcal{M}'$ which simulates $\mathcal{M}$ for
              $3n\cdot c$ steps where $c$ is the number of configurations of
              $M$
            \item Output $e = (e_{1},\dots,e_{k})^{\ast}$
            \item Assume $e$ is not universal; then it loops in $V'$ with
              $|V'|<n$
            \item $\mathcal{M}$ has to enter a vertex twice with the same
              configuration
            \item Contradicts universality of the exploration sequence of
              $\mathcal{M}$
          \end{itemize}
        }
      }
  \end{itemize}
  \pause
  \pause
\end{frame}

\begin{frame}
  \frametitle{3-Regularity}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Establish 3-regularity
        \begin{itemize}
          \item Recall: \parbox{3\baselineskip}{\resizebox{!}
            {3\baselineskip}{\input{tikz/blowup}}}
          \item<2-> Transform $v$ to $3\cdot d_{v}$ distinct vertices
        \end{itemize}
    \end{itemize}
    \column{0.5\textwidth}
    \uncover<3->{\resizebox{\textwidth}{!}{\input{tikz/regsim}}}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Exploring pebble machine}
  \begin{itemize}
    \item Graph $G$
    \item Transform to 3-regular $G_{\text{reg}}$
    \item Applying PES-Lemma to $G_{\text{reg}}$ yields:
  \end{itemize}
  \begin{theorem}
    There exists a $(\mathcal{O}(1), 0, \mathcal{O}(\log n))$-pebble machine
    that moves along a closed walk and either explores the graph or visits at
    least $n$ distinct vertices, for any graph with bounded degree.
  \end{theorem}
\end{frame}

\section{Pebble simulation}
\begin{frame}
  \begin{theorem}
    There is a constant $c\in\mathcal{N}$, such that for any Graph $G$
    with bounded degree and any $(s,p,2m)$ pebble machine $\mathcal{M}$, there
    exists a $(cs,p+c,m)$ pebble machine $\mathcal{M}'$ that simulates the walk
    of $\mathcal{M}$ or explores $G$.
  \end{theorem}
\end{frame}

\begin{frame}
  \frametitle{Tape simulation via pebbles}
  \begin{itemize}
    \item Simulate tape of size $m$ by pebbles:
      \begin{itemize}
        \item<3-> Separate into memory blocks of size $m_{1}$
        \item<4-> Represent every block by one of $\left\{p_{0},\dots,p_{\frac{m}{m_{1}}-1}\right\}$ pebbles
      \end{itemize}
  \end{itemize}
  \uncover<2->{\input{tikz/tapesimint}}
\end{frame}

@@

\begin{frame}
  \frametitle{Transfer tape-content through movements}
  \begin{itemize}
    \item Simulated machine moves
    \item Preserve tape-content:
      \begin{itemize}
        \item Drop $p_{\text{next}}$ on next vertex
        \item Transfer encoding pebbles onto walk $\omega'$ starting from
          next vertex
        \item Preserve \enquote{state} by placing at same index
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Recall: Mechanisms of simulation}
  \begin{itemize}
    \item Simulation of $M_{\text{Walk}}$ which explores at least $2^{m_{1}}$
      distinct vertices \uncover<2->{\checkmark}
    \item Identify \emph{distinct} vertices of the walk \uncover<3->{\checkmark}
    \item Read from and write to simulated tape \uncover<4->{\checkmark}
    \item Preserve tape content through steps of simulated agent \uncover<5->{\checkmark}
  \end{itemize}
  \uncover<6->{
    \begin{theorem}
      There is a constant $c\in\mathcal{N}$, such that for any Graph $G$
      with bounded degree and any $(s,p,2m)$ pebble machine $\mathcal{M}$, there
      exists a $(cs,p+c,m)$ pebble machine $\mathcal{M}'$ that simulates the walk
      of $\mathcal{M}$ or explores $G$.
    \end{theorem}
  }
\end{frame}

\begin{frame}
  \frametitle{Exploring graphs with $\log\log n$ pebbles}
  \begin{itemize}
    \item There is a $(c,0,c\cdot\log n)$ pebble machine which explores
      a graph $G$ with $n$ vertices
    \item Applying above simulation $\log \log n$ times yields
      $(c'^{\log\log n}\cdot c, c'\cdot\log\log n, c)$ pebble machine
      \begin{theorem}
        Any bounded-degree graph on at most $n$ vertices can be explored using
        $\mathcal{O}(\log\log n)$ pebbles and memory.
      \end{theorem}
  \end{itemize}
\end{frame}
